\documentclass{article}
\usepackage[]{graphicx}
\usepackage[]{color}
\usepackage{hyperref}
\usepackage{listings}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
  language=bash,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{red},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
}

% maxwidth is the original width if it is less than linewidth
% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

\title{Getting Started with Raspberry Pi}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}


\author{Anna Burns, Kyle McCarty, and Marc Los Huertos}
\title{Getting Started with Raspberry Pi}

\begin{document}

\maketitle

\newpage

\tableofcontents

\newpage

\section{Introduction}

\subsection{What is a Raspberry Pi?}

The Raspberry Pi is an tiny computer, that includes a microprocessor, a bit of memory, a slot for an SD card, input/output (I/O) ports, e.g. HDMI, USB, headphone, camera, and some general purpose input/output (GPIO) pins for various types of electrical connectors.

\subsection{Why use the Raspberry Pi?}

Generally, Raspberry Pis draw considerably less power than regular computers, are a lot smaller, and are relatively cost-effective. In addition, the GPIO pins allow for connecting and controlling various types of electrical components, such as LEDs and sensors. Raspberry Pis are very flexible devices. They can be used for personal computers, home survaillance systems, weather stations, adblockers for your home network, retro gaming machines, as an AI assistant, and so much more! In this class, we'll be using it as an environmental monitoring device.

\section{Unpacking and Connecting the Pi}

\subsection{Packaging List}

\subsubsection{``Vilros RP Zero W Basics Kit''}

\begin{enumerate}
  \item Raspberry Pi Zero W board
  \item Case, with 3 covers
  \item 2.5A power supply
  \item Heatsink
  \item HDMI to mini-HDMI adapter
  \item USB to micro-USB adapter
  \item Header pin diagram
  \item Camera module adapter (not used)
\end{enumerate}

\subsubsection{Other items}

\begin{enumerate}
  \item SD card
  \item SD card to USB adapter
  \item USB multiport adapter
  \item Breadboard
  \item Wires
  \item Line level converter (LLC)
  \item MCP3008
  \item MQ-135
  \item MQ-XXX
  \item MQ-XXX
\end{enumerate}

\subsection{Install Raspberry Pi OS on SD card}

\begin{enumerate}
  \item Download Raspberry Pi Imager for your main computer's operating system (OS) at \url{https://www.raspberrypi.org/downloads/}
  \item Install Raspberry Pi Imager
  \item Connect the Pi's SD card to your main computer.
  \item Use Raspberry Pi Imager to install/write Raspberry Pi OS to SD card.
  \item Alternatively, manually copy Raspberry Pi OS and NOOBS to SD card, using the link above.
\end{enumerate}

\subsection{Add network configuration file to SD card for a secure shell (SSH) connection.} \label{networkssh}

\begin{enumerate}
  \item Add a ``ssh'' file to your boot partition on the SD card.
  \begin{itemize}
  \item Do this by creating a text file named \textbf{``ssh''}.
  \item Make sure the file you create has no file extension.
  \end{itemize}
  \item Create and add a file called \textbf{``wpa\_supplicant.conf''} to the boot partition on the SD card. This can be created with any text editor.
  \begin{itemize}
  \item Make sure the file you created has the \textbf{``.conf''} file extension in the name.
  \item The \textbf{``wpa\_supplicant.conf''} needs to have the WiFi network information in it for the Raspberry Pi to connect on boot up.
  \item Modify this file with a text editor and include the following information:
  \end{itemize}
\newpage  
\begin{lstlisting}
ctrl_interface=DIR=/var/run/wpa_supplicant GROUP=netdev
update_config=1
country=US

network={
 ssid="WIFI NETWORK NAME"
 psk="WIFI PASSWORD"
}
\end{lstlisting}

  \begin{itemize}
  \item If you are not in the United States, input your country's ISO code instead of \textbf{``US''} on the \textbf{\textit{``country=US''}} line.
  \item Make sure to change \textbf{``WIFI NETWORK NAME''} to your WiFi network name.
  \item Make sure to change \textbf{``WIFI\_PASSWORD''} to your WiFi network's password.
  \end{itemize}
  
\end{enumerate}

\subsection{Assemble and Connecting the Pi}

\subsubsection{Putting the Pi together}

\begin{enumerate}
  \item Safely eject the SD card from your main computer.
  \item Place the SD card in the Raspberry Pi Zero W's SD card slot.
  \item Attach the Pi to the bottom Vilros case making sure to line the dowels within the case with the mouting holes on the Pi.
  \item \textbf{Do not} attached the top part of the case yet. Being able to see the Pi's board will help you with pin determination.
  \item Attach your peripheral devices, \textit{if you have them}. This includes a monitor, mouse, and keyboard. It is okay to not have them.
  \begin{itemize}
    \item You will need to use the HDMI to micro-HDMI adapter to hook a monitor up to the Raspberry Pi Zero W.
    \item Also, you'll need to use the Anker multiport USB adapter to use a keyboard \textbf{and} a mouse. The Raspberry Pi Zero W only has one micro-usb port.
    \begin{itemize}
      \item You can get away with using just a monitor and keyboard, but if you are not comfortable navigating with a keyboad only, it'll be difficult.
    \end{itemize}
  \end{itemize}
  \item Lastly, making sure the the power adapter for the Pi is plugged in to a power source, and verifying the switch on the adapter is in the \textbf{``off''} position, connect the power cable into the Pi.
  \item Now, turn the power switch on.

\textbf{NOTE!} Make sure not to turn the device off, or on/off/on, otherwise the \textbf{``ssh''} and \textbf{``wpa\_supplicant.conf''} files will not be in your boot directory on the SD card anymore. If, for whatever reason, the Pi loses power after putting those file in the boot directory, you will have to go back and add them, as in step \textbf{\ref{networkssh}}. This is \textbf{especially} important for \textbf{\textit{headless}} users (users with no monitor).

\textbf{NOTE!} This is where \textbf{\textit{headless}} users will have to remotely connect to the Pi, while people who have monitors, mice, and keyboards won't have to remotely connect. For those who aren't connecting remotely, skip ahead to step \textbf{\ref{updateupgrade}}

\end{enumerate}

\section{Accessing and updating Raspberry Pi OS}

Now, there are different ways to remotely connect with the Raspberry Pi if you are a \textbf{\textit{headless}} user. Unfortunately though, you will have to connect to the Pi using \textbf{SSH} the \textbf{first time}. After that, you can install a remote server on the Pi so you can connect with it using a program with a GUI. I recommend using SSH, though, as it is a \textbf{\textit{much}} quicker connection.

\subsection{Remote connection via SSH}

\subsubsection{Installing Raspberry Pi Finder and finding Pi's IP address}

The goal with this step is to find the \textbf{local IP address} of the Pi. There are a lot of different ways to do this. If you are computer savvy, go ahead and find the IP address of your Pi and ingnore this step.

\begin{enumerate}
  \item The easiest way to find the IP address of your Pi, if you don't know networking or computers that well, is to use the \textbf{Raspberry Pi Finder} by Adafruit.
  \item With your main computer, go to this websiteto download the application for your OS: \newline \url{https://github.com/adafruit/Adafruit-Pi-Finder/}
  \item Scroll down and click the link that says \textbf{``Download the latest release''}
  \item Scroll down to \textbf{``Assets''} and there you will see the .zip files of the program for the different OSs.
  \begin{itemize}
    \item \textbf{``osx''} is for Mac users.
    \item \textbf{``win32''} is for Windows user.
  \end{itemize}
  \item Download the .zip file for your system and then unzip it when finished downloading.
  \item Run Raspberry Pi finder.
  \item Click \textbf{``Find my Pi!''} for the program to locate your Pi. Wait a few minutes if it doesn't find it immediately. Sometimes it can take quite a while!
  \item The IP address should be listed when finished. It should look something like \textbf{``192.168.1.XXX''}. \label{ipaddress}
  \item Use the IP address to connect to the Pi.
\end{enumerate}

\subsubsection{SSH connection}
\label{ssh}

\begin{itemize}
    \item For Windows users, you will want to use the \textbf{Command Prompt}.
    \begin{enumerate}
      \item Press the \textbf{\textit{Windows}} key on your keyboard, type \textbf{``CMD''}, and \textbf{Enter}.
      \item In the Command Prompt, type the command below making sure to substitute ``192.168.1.XXX'' with the IP address found in step \ref{ipaddress}
      \begin{lstlisting}
      ssh pi@192.168.1.XXX 
      \end{lstlisting}
    \end{enumerate}
    \item For Mac users, you will want to use the \textbf{\textit{Terminal}}.
    \begin{enumerate}
      \item Press the \textbf{\textit{Command + Spacebar}} keys on your keyboard, type \textbf{``Terminal''}, and \textbf{Enter}.
      \item In Terminal, type the command below making sure to substitute ``192.168.1.XXX'' with the IP address found in step \ref{ipaddress}
      \begin{lstlisting}
      ssh pi@192.168.1.XXX 
      \end{lstlisting}
    \end{enumerate}
\end{itemize}
\begin{enumerate}
  \item For both Mac and Windows users, you should now see a prompt in the command line asking for the Pi's password. 
  \newline 
  \newline 
  By default, the password is \textbf{``raspberry''}. 
  \newline 
  \newline 
  The command line won't show the password while you are typing it, nor will it show placeholder characters. You will need to just type it and \textbf{Enter}.
  \item If all went well, you should see that you've connected to your Pi via the command line interface (CLI). You should see the text: 
  \begin{lstlisting}
  pi@raspberrypi/:~ $_ 
  \end{lstlisting}
  \item There will also be a warning that the Pi's password hasn't been changed. We will want to change that, but we are not going to do that right now.
\end{enumerate}

\subsection{Update and Upgrading Raspberry Pi OS} 
\label{updateupgrade}

Here we will be updating the Raspberry Pi. This is the first thing we are going to want to do to make sure there are now errors later on. We will need to implement a few commands into the Pi's command line. Don't be afraid! Its easy!

\begin{enumerate}
  \item First we are going to want the Pi to talk to the official Raspberry Pi servers to get the latest version list of our programs on the Pi. In your SSH session, type the following command:
  \begin{lstlisting} 
  sudo apt update
  \end{lstlisting}
  \item Secondly, we want the Pi to compare the version list with its current packages and programs and update where needed. In your SSH session, type:
  \begin{lstlisting}
  sudo apt full-upgrade
  \end{lstlisting}
  \item You will be asked if you are sure you want to upgrade. Type \textbf{``y''} and \textbf{Enter}.
  \item Once it is finished, it will show this once again:
  \begin{lstlisting}
  pi@raspberrypi/:~ $_ 
  \end{lstlisting}
  \item Congratulations! Your Pi has been upgraded!
\end{enumerate}

\section{Configuring the Pi}

\begin{itemize}
  \item Hopefully you still have your current SSH connection going with your Pi. If not, try to create one. If you exited the Command Prompt/Terminal, and severed the SSH connection, just reopen them and reconnect via SSH as in step \ref{ssh}.
  \item If \textbf{power} was severed from the Pi, you will need to repeat section \ref{networkssh} over again. The Pi discards these files after it boots up, so if you shut off the Pi, or lose power, you'll have to re-add them to the boot directory.
\end{itemize}

\subsection{Changing the Pi User Password}

As a security measure, since your Pi is in your WiFi network, you'll want to change the password for the Pi.

\begin{enumerate}
  \item Run the Raspberry Pi configuration utility by using this command in the CLI:
  \begin{lstlisting}
  sudo raspi-config
  \end{lstlisting}
  \item Navigate to \textbf{``Change User password for the 'pi' user''} and hit \textbf{Enter}.
  \item It will prompt you that it is going to ask for the new password. Press \textbf{Enter}, and type your new password followed by \textbf{Enter} again.
  \item Verify the password by typing it again.
  \item Your Pi now has your new password. Don't forget it!
\end{enumerate}

\subsection{Network Options}

We want to make sure the Pi's network options are configured properly so we don't have to add a \textbf{wpa\_supplicant.conf} file to the boot directory when booting anymore.

\begin{enumerate}
  \item While still in the raspi-config utility, navigate to \textbf{``Network Options''}.
  \item Navigate to \textbf{``Wireless LAN''} and hit \textbf{Enter}.
  \item It will prompt you to enter your WiFi \textbf{SSID}. This is the name of the WiFi network. Enter it and press \textbf{Enter}.
  \item Next it will ask for the WiFi password/passphrase. Type that in and hit \textbf{Enter}.
  \item If done correctly, it will notify you it was successfully done via the CLI.
  \item Your Pi now has the wireless network configured properly.
\end{enumerate}

\subsection{Boot Options}

This next configuration is more for fun and is \textbf{OPTIONAL}. We are going to make the Raspberry Pi \textbf{``text boot''} instead of just showing a splash screen. This is fun because it shows you the boot process via the CLI so you can see \textbf{\textit{exactly}} what is starting during boot up!

\begin{enumerate}
  \item While still in the raspi-config utility, navigate to \textbf{``Boot Options''}.
  \item Navigate to \textbf{``Splash Screen''}.
  \item It will prompt you if you want to see the \textbf{splash screen} at boot. Select \textbf{``NO''}
  \item It will then tell you that the splash screen at boot is disabled. Woohoo!
\end{enumerate}

\subsection{Localization Options}

This may or may not be a problem for some people. I had to make sure my time zone was set correctly. I believe by default the Pi's time zone is set for Europe.

\begin{enumerate}
  \item While still in the raspi-config utility, navigate to \textbf{``Localisation Options''}.
  \item Navigate to \textbf{``Change Time Zone''}.
  \item In a few seconds it will prompt you to choose your \textbf{geographic area}. Select whichever corresponds to yours. In my case it was the US.
  \item It will then tell ask you what  your \textbf{time zone} is. Select what applies. In my case is was Pacific Ocean.
  \item Then, it will flash the CLI with the Coordinated Universal Time (UTC) and your time zone's time and go back to the configuration utility.
\end{enumerate}

\subsection{Interfacing Options (enable SSH)}

\textbf{IMPORTANT} We need to make sure we enable SSH in the Pi's configuration. This is extremely important and allows us to connect to the Pi via SSH in all future instances without having to add a ssh file to the boot directory.

\begin{enumerate}
  \item While still in the raspi-config utility, navigate to \textbf{``Interfacing Options''}. Remember where this is located because we will need to access this later on.
  \item Navigate to \textbf{``SSH''}.
  \item The Pi will then ask if you want the \textbf{SSH server} enabled. Select \textbf{Yes}.
  \item Wait a few seconds and the Pi should notify you the SSH server is enabled.
\end{enumerate}

\subsection{Virtual network Computing (VNC) Option}

\textbf{NOTE} For those who find the CLI a bit tricky, or just don't like it, you can enable a virtual network computing software called RealVNC before shutting down the Pi. This will let you control the Pi via a graphical user interface (GUI), like what people are normally used to on MacOS or Windows.

\begin{enumerate}
  \item While still in the raspi-config utility, navigate to \textbf{``Interfacing Options''}.
  \item Navigate to \textbf{``VNC''} and select it.
  \item The Pi will prompt you if you want the \textbf{VNC Server} enabled. Select \textbf{``Yes''}.
  \item Wait a second and the Pi should let you know that the \textbf{VNC Server} is now enabled. This server will start automatically when the Pi boots up.
  \item You still need to download \textbf{RealVNC} \textbf{\textit{Viewer}} on the computer you want to remotely access the Pi with.
  \item \textbf{RealVNC Viewer} can be downloaded at:
  \newline
  \url{https://www.realvnc.com/en/connect/download/viewer/}
  \item Download the client that is for your OS and then install it.  
\end{enumerate}

\subsection{Finishing Up}

You Pi should be configured now! When the Pi does get turned off and turned back on, it should boot and connect to your WiFi and allow you to SSH into it. Now, we still need to close the Pi's configuration utility and shut down the Pi.

\begin{enumerate}
  \item While still in the raspi-config utility, navigate down as far as it lets you \textbf{(About raspi-config)}.
  \item You will have to press the right arrow key to get down to \textbf{``FINISH''}. Tricky and a little unintuitive, I know...
  \item Select \textbf{FINISH} and it should exit the utility back to the CLI.
  \item Next, run this last command in the CLI to shutdown the Pi:
  \begin{lstlisting}
  sudo shutdown -h now
  \end{lstlisting}
  \item Soon the Pi will shutdown and you'll get a message in your Command Prompt/Terminal something along the lines that the connection was closed or lost.
  \item Flip the switch on the power supply to the Pi to make sure power is not going to the Pi.
\end{enumerate}

\textbf{CONGRATULATIONS}! You done configuring the Pi. \textit{Whew}...

\end{document}
